### 2015-04-02 paper/abstract.tex
Updated paper. Separated out sections and added CP figure.

\begin{abstract}
\begin{comment}
SPL (+CPL)

WTD

ROF

IMP
\end{comment}
\end{abstract}