### 2015-04-02 paper/futurework.tex

\section{Future Work}

In addition to that already mentioned: there remains much to be done.