### 2015-04-02 paper/paper.tex

\documentclass{mpaper}

\usepackage{comment}
\usepackage{lstform}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{cmll}
\usepackage{prooftree}
\usepackage{amsmath}
\newcommand{\typestate}[2]{{#1}\gg{#2}}
\usepackage[style=alphabetic,sorting=nymt,natbib=true,
  maxbibnames=100]{biblatex}
\addbibresource{paper.bib}

\DeclareSortingScheme{nymt}{
  \sort{
    \field{author}
    \field{year}
    \field{month}
    \field{title}
  }
}

%\input{cpgv}
\input{propmacros}

\newcommand{\fpop}{System F${}^\circ$\xspace}

\begin{document}

\title{Propositions as Sessions, Mechanically}
\author{Craig McLaughlin}
\matricnum{1002524}

\maketitle

\input{abstract}
\input{introduction}
\input{background}
\input{formalisation}
\input{futurework}
\input{conclusion}
\input{acknowledgements}

\sloppy
\printbibliography

\end{document}