### 2015-04-03 paper/futurework.tex
Updated paper: CP section and added figures.

\section{Future Work}

\citeauthor{??} extend \citeauthor{Wadler:2014}'s work to support polymorphism
in GV. (polymorphism in CP) Future work is to support reduction under these
terms.