### 2015-04-14 paper/conclusion.tex
Added conclusion.

\section{Conclusion}

The current paper presented work towards a formalisation of
\citeauthor{Wadler:2014}'s two-tier GV/CP system. The Coq development defines
the process calculus CP and the functional language GV using state-of-the-art
techniques from the programming language and interactivee theorem proving
communities. Proof of subject reduction for CP has been formalised along with
work towards proving sound both cut elimination and the GV to CP CPS
translation. While ultimately there remains work to be done to complete the
development, several issues and guidelines have been identified to be
considered by future endeavours in the area. Finally, some ideas for future
work, on GV/CP specifically but also generally in the area of session-based
type system formalisations, were proffered.